% Theseus pitch deck — "A philosopher in a box."
% Compiled with pdflatex via ./build_deck.sh.
%
% Slide 11 ("What's working") pulls live numbers from the DB through
% live_snapshot.py which writes slide11_data.tex. If the snapshot
% script cannot read the live DB the build aborts loudly rather than
% ship fake numbers.

\documentclass[10pt]{article}
\input{_preamble.tex}

\begin{document}

% --------------------------------------------------------------------
% 1. Title
% --------------------------------------------------------------------
\begin{slide}{1}{Title}
  \vspace*{1.6in}
  \begin{center}
    {\rmfamily\bfseries\fontsize{56pt}{60pt}\selectfont\color{ink}
      A philosopher in a box.}\\[0.5in]
    {\sffamily\fontsize{14pt}{18pt}\selectfont\color{amber}
      THESEUS\quad\textbullet\quad 2026}
  \end{center}
\end{slide}

% --------------------------------------------------------------------
% 2. The problem
% --------------------------------------------------------------------
\begin{slide}{2}{Problem}
  \slidehead{Quants solved one layer.}
  \slidesub{The next layer up is unworked.}
  \sliderule

  Quantitative firms abstract \emph{numerical patterns} out of price
  data and arbitrage them at scale. The architecture is well
  understood: data archive, model trainer, fitted parameters, live
  signal, position.

  \vspace{0.2in}

  Above the price layer is the \emph{logic layer} --- the patterns
  inside arguments themselves, the principles that organise how a
  reader maps a text to a position. No firm has industrialised that
  layer. The reading is done by hand, the principles never leave the
  reader's head, and the bet, when it comes, is a one-off.

  \vspace{0.25in}

  \callout{There is room for one Renaissance up the stack.}
\end{slide}

% --------------------------------------------------------------------
% 3. The thesis
% --------------------------------------------------------------------
\begin{slide}{3}{Thesis}
  \slidehead{Principles are arbitrageable.}
  \slidesub{The same machine, one level up.}
  \sliderule

  A principle is a logical pattern abstracted from a text --- the
  stable shape underneath an argument, expressible in a sentence or
  two, applicable to situations the original author never saw.

  \vspace{0.2in}

  Theseus reads a curated corpus the way Renaissance reads price
  tape. The synthesizer is the trainer; the principle is the fitted
  artefact; the algorithm is the live executor; the conclusion is the
  position. The bet, when the algorithm fires sharply enough, is the
  output of the same machine --- operating on logic instead of
  numbers.

  \vspace{0.2in}

  \callout{We are the Renaissance Technologies of formal logic.}
\end{slide}

% --------------------------------------------------------------------
% 4. The machine, abstract
% --------------------------------------------------------------------
\begin{slide}{4}{The machine — abstract}
  \slidehead{Inputs $\rightarrow$ engine $\rightarrow$ conclusions $\rightarrow$ bets.}
  \sliderule

  \vspace{0.15in}

  \begin{center}
  \begin{tikzpicture}[node distance=1.3cm and 1.5cm,
      every node/.style={font=\sffamily\bfseries\fontsize{11pt}{13pt}\selectfont,
                         draw=amber,line width=0.6pt,
                         rounded corners=2pt,inner sep=10pt,
                         text=ink,fill=parchment},
      arrow/.style={-{Latex[length=2.5mm]},line width=0.8pt,color=amber}]
    \node (inputs) {INPUTS};
    \node[right=of inputs] (engine) {SYNTHESIZER ENGINE};
    \node[right=of engine] (concl) {CONCLUSIONS};
    \node[right=of concl] (bets) {BETS};
    \draw[arrow] (inputs) -- (engine);
    \draw[arrow] (engine) -- (concl);
    \draw[arrow] (concl) -- (bets);
  \end{tikzpicture}
  \end{center}

  \vspace{0.3in}

  Inputs are the curated corpus and live observations of the world.
  The synthesizer extracts principles. Conclusions are principles
  applied to live observations. The bet is the conclusion expressed
  as a position. Every arrow is recorded, replayable, and gradable.
\end{slide}

% --------------------------------------------------------------------
% 5. The machine, concrete
% --------------------------------------------------------------------
\begin{slide}{5}{The machine — concrete}
  \slidehead{The Theseus pipeline.}
  \sliderule

  \begin{center}
  \begin{tikzpicture}[node distance=0.45cm and 0.85cm,
      every node/.style={font=\sffamily\fontsize{9pt}{11pt}\selectfont,
                         draw=amber,line width=0.5pt,
                         rounded corners=2pt,inner sep=7pt,
                         text=ink,fill=parchment,minimum width=1.55in,
                         align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.7pt,color=amber}]
    \node (corpus) {Curated corpus};
    \node[right=of corpus] (synth) {Synthesizer};
    \node[right=of synth] (princ) {Principles};
    \node[right=of princ] (algo) {Algorithms};
    \node[below=of corpus] (live) {Live observations\\ (Currents)};
    \node[below=of algo] (conc) {Conclusions};
    \node[below=of conc] (memo) {Memos};
    \node[left=of memo] (port) {Portfolio agent};
    \node[left=of port] (bet) {Bets\\ (Forecasts / Equities)};
    \draw[arrow] (corpus) -- (synth);
    \draw[arrow] (synth) -- (princ);
    \draw[arrow] (princ) -- (algo);
    \draw[arrow] (live) -| (algo);
    \draw[arrow] (algo) -- (conc);
    \draw[arrow] (conc) -- (memo);
    \draw[arrow] (memo) -- (port);
    \draw[arrow] (port) -- (bet);
  \end{tikzpicture}
  \end{center}

  \vspace{0.1in}

  \slidesub{Every named box is a surface on the public site.}
\end{slide}

% --------------------------------------------------------------------
% 6. Example — arms race
% --------------------------------------------------------------------
\begin{slide}{6}{Example I}
  \slidehead{The arms-race algorithm.}
  \slidesub{One principle, one observation, one bet.}
  \sliderule

  \begin{stack}{Principle invoked}{Live observation}
    Two-party arms races, when one party signals a credible escalation
    and the other is institutionally constrained from matching it,
    converge on a settlement on terms favourable to the unconstrained
    party --- not on the open conflict that surface news coverage
    predicts. &
    A credible escalation signal from party A; documented institutional
    constraints on party B (legislative, fiscal, alliance-internal);
    public commentary still pricing open conflict. \\
  \end{stack}

  \vspace{0.3in}

  \begin{stack}{Algorithm output}{Implied bet}
    Conclusion: settlement on A-favourable terms within the horizon,
    not open conflict. Confidence band documented; conditions for
    revision named. &
    Long the prediction-market contract for settlement-by-date;
    short the contract for open conflict in the same window. Memo
    written; portfolio agent sized; bet recorded. \\
  \end{stack}
\end{slide}

% --------------------------------------------------------------------
% 7. Example — onion futures (Strauss frame)
% --------------------------------------------------------------------
\begin{slide}{7}{Example II}
  \slidehead{Onion futures.}
  \slidesub{The quant sees price oscillation. We read why.}
  \sliderule

  Onion futures were banned in 1958 after a corner that ricocheted
  through the price tape for years. A quant looking at the chart sees
  amplitude, volatility, a regime change. A reader who has worked
  through the Senate testimony, the speculators' position book, and
  the producers' political reach sees \emph{the logic of the
  ban itself} --- the principle that produced the price action.

  \vspace{0.2in}

  \callout{``To write between the lines.''}
  {\sffamily\fontsize{9pt}{11pt}\selectfont\color{amberdim}%
   --- Leo Strauss, Persecution and the Art of Writing (1952)}

  \vspace{0.25in}

  The Strauss frame is operational, not decorative. The synthesizer
  is built to read both the surface argument and the constraint that
  shaped it. A principle library that holds those constraints
  predicts the price action that the chart only describes after the
  fact.
\end{slide}

% --------------------------------------------------------------------
% 8. Renaissance comparison
% --------------------------------------------------------------------
\begin{slide}{8}{Renaissance comparison}
  \slidehead{The same stack, one layer up.}
  \sliderule

  \begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.3\textwidth}
                    @{\hspace{0.04\textwidth}}>{\raggedright\arraybackslash}p{0.3\textwidth}
                    @{\hspace{0.04\textwidth}}>{\raggedright\arraybackslash}p{0.3\textwidth}@{}}
    \sffamily\bfseries\color{amber}LAYER &
    \sffamily\bfseries\color{amber}RENAISSANCE TECH &
    \sffamily\bfseries\color{amber}THESEUS \\
    \midrule
    Input archive & Tick data, fundamentals, alternative data & Curated corpus of texts (memos, papers, talks, filings) \\[0.15in]
    Trainer & Statistical model fitting & Synthesizer (principle extraction) \\[0.15in]
    Fitted artefact & Parameters of an alpha model & Principle (logical pattern, expressible in a sentence) \\[0.15in]
    Live executor & Strategy code on live ticks & Algorithm on live observations \\[0.15in]
    Output & Position & Conclusion $\rightarrow$ memo $\rightarrow$ bet \\[0.15in]
    Edge & Proprietary models and execution & Proprietary corpus, principles, and algorithm registry \\
  \end{tabular}
\end{slide}

% --------------------------------------------------------------------
% 9. Bet domains
% --------------------------------------------------------------------
\begin{slide}{9}{Bet domains}
  \slidehead{Polymorphic across where logic can be priced.}
  \sliderule

  The machine is bet-domain-polymorphic. The same principle library,
  the same algorithm registry, the same portfolio agent --- different
  venues:

  \vspace{0.2in}

  \begin{itemize}[leftmargin=0.35in,itemsep=0.12in,topsep=0pt]
    \item \textbf{Equities.} Theses with multi-quarter horizons,
          executed through the firm's brokerage relationships.
    \item \textbf{Prediction markets.} Discrete questions with a
          settlement date; the calibration backbone of the system.
    \item \textbf{Advisory.} The same conclusions, sold as judgement
          to clients who pay for it directly.
    \item \textbf{Scientific.} Hypotheses the principle library
          predicts; published, falsified, or extended.
    \item \textbf{Private markets (eventually).} Once the calibration
          record across the public venues is long enough to justify
          the cycle time.
  \end{itemize}
\end{slide}

% --------------------------------------------------------------------
% 10. The team
% --------------------------------------------------------------------
\begin{slide}{10}{Team}
  \slidehead{Team.}
  \sliderule

  \IfFileExists{team_data.tex}{\input{team_data.tex}}{%
    \slidesub{Team data is generated at build time from the live
    founder roster.}

    The build script writes \texttt{team\_data.tex} from the same
    \texttt{founder} table the public About page reads from. If the
    build cannot reach the database, the slide falls back to this
    note --- but the founder copy in the deck PDF that ships should
    always be the generated version.
  }
\end{slide}

% --------------------------------------------------------------------
% 11. What's working — live snapshot
% --------------------------------------------------------------------
\begin{slide}{11}{What's working}
  \slidehead{What's working.}
  \slidesub{Numbers pulled at build time. The deck is never stale.}
  \sliderule

  \input{slide11_data.tex}

  \vspace{0.2in}

  {\sffamily\fontsize{8pt}{10pt}\selectfont\color{amberdim}%
   Snapshot generated by \texttt{live\_snapshot.py}. If the live DB is
   unreachable, the build fails loudly rather than ship a stale or
   fabricated snapshot.}
\end{slide}

% --------------------------------------------------------------------
% 12. The ask
% --------------------------------------------------------------------
\begin{slide}{12}{The ask}
  \slidehead{The ask.}
  \sliderule

  \begin{stack}{Capital}{Time horizon}
    \texttt{\$\_\_\_\_\_\_\_\_\_\_\_} &
    \texttt{\_\_\_\_\_\_ months / years} \\
  \end{stack}

  \vspace{0.3in}

  \begin{stack}{Structure}{Use of proceeds}
    \texttt{\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_} &
    \texttt{\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_} \\
  \end{stack}

  \vspace{0.3in}

  \slidesub{Template. The founder fills the four blanks before each
  conversation; the rest of the deck is identical across pitches.}
\end{slide}

% --------------------------------------------------------------------
% 13. What we are NOT
% --------------------------------------------------------------------
\begin{slide}{13}{What we are NOT}
  \slidehead{What we are not.}
  \sliderule

  \callout{Theseus is not a SaaS product. The reasoning
  architecture is our edge.}

  \vspace{0.3in}

  A SaaS distribution would force a single workflow on every buyer
  and would expose the corpus, the principle library, and the
  algorithm internals to anyone with a credit card. The machine is
  valuable in proportion to how much of the firm's specific reading
  has been encoded into it; selling shells of that would commodify
  the surface and erode the edge.

  \vspace{0.2in}

  We invest with the machine. We do not sell access to it. The
  public surfaces --- Currents, Forecasts, Memos, Algorithms,
  Principles, Knowledge graph --- are the firm thinking in public.
  They are not the product.
\end{slide}

% --------------------------------------------------------------------
% 14. Contact + appendix pointer
% --------------------------------------------------------------------
\begin{slide}{14}{Contact}
  \slidehead{Contact.}
  \sliderule

  \begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.4\textwidth}
                    @{\hspace{0.04\textwidth}}>{\raggedright\arraybackslash}p{0.5\textwidth}@{}}
    \sffamily\bfseries\color{amber}Firm &
      \href{https://www.theseuscodex.com}{theseuscodex.com} \\[0.15in]
    \sffamily\bfseries\color{amber}Email &
      \href{mailto:hello@theseuscodex.com}{hello@theseuscodex.com} \\[0.15in]
    \sffamily\bfseries\color{amber}About &
      \href{https://www.theseuscodex.com/about}{theseuscodex.com/about} \\[0.15in]
    \sffamily\bfseries\color{amber}Memos &
      \href{https://www.theseuscodex.com/memos}{theseuscodex.com/memos} \\[0.15in]
    \sffamily\bfseries\color{amber}Algorithms &
      \href{https://www.theseuscodex.com/algorithms}{theseuscodex.com/algorithms} \\[0.15in]
    \sffamily\bfseries\color{amber}Three-minute script &
      \texttt{docs/pitch/three\_minute\_script.md} \\
  \end{tabular}

  \vspace{0.4in}

  \callout{The machine is our edge.}
\end{slide}

\end{document}
