\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{titlesec}
\usepackage[hidelinks,colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black]{hyperref}
\usepackage{xcolor}
\usepackage{fancyhdr}

\titleformat{\section}{\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries}{\thesubsection}{1em}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Theseus Auto-Paper --- adversarial-audit-2a74eb9b}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\setlength{\parindent}{0pt}

\newcommand{\todomark}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\rowref}[1]{\texttt{\small[row:#1]}}

\title{\textbf{Adversarial probing of hidden assumptions: a firm cluster}\\[6pt]\large Cluster adversarial-audit-2a74eb9b}
\date{May 2026}
\author{Theseus Auto-Paper Generator}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{center}
\fbox{%
  \begin{minipage}{0.92\textwidth}
    \centering\small
    \textbf{Disclosure:} machine-drafted, founder-reviewed.\\
    Lead conclusion: \texttt{conclusion-adversarial-audit-1}.\\
    Methodology root: \texttt{profile-adversarial-audit-1}
    (adversarial-audit).
  \end{minipage}%
}
\end{center}

\vspace{1.5em}

\section*{Abstract}
This paper distills 2 firm conclusion(s) anchored on conclusion-adversarial-audit-1, all sharing the methodology root "Adversarial probing of hidden assumptions". The cluster's claims are stress-tested against 1 resolved forecast(s) drawn from the firm's calibration ledger. Lead claim: Adversarial review surfaces the hidden assumption a friendly read leaves buried; we will block publication if an objection is unaddressed.

\section{Introduction}
The Theseus engine routinely produces firm conclusions, attaches them to a methodology profile, and tracks how those conclusions perform when their predictive consequences are settled in public forecast markets. This paper presents one such cluster of 2 conclusion(s), drawn together by their shared methodology root and by at least one resolved forecast that touches the cluster. The cluster's lead claim states: "Adversarial review surfaces the hidden assumption a friendly read leaves buried; we will block publication if an objection is unaddressed.".

\subsection*{Conclusions in this cluster}
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{conclusion-adversarial-audit-1} \rowref{conclusion:conclusion-adversarial-audit-1} --- Adversarial review surfaces the hidden assumption a friendly read leaves buried; we will block publication if an objection is unaddressed.
  \item \textbf{conclusion-adversarial-audit-2} \rowref{conclusion:conclusion-adversarial-audit-2} --- A buried assumption stays buried under a friendly read, so the firm runs the adversarial reviewer first.
\end{itemize}

\section{Methods}
The cluster shares a single methodology root,
\textit{Adversarial probing of hidden assumptions} (\texttt{profile-adversarial-audit-1}, pattern \emph{adversarial-audit}).

\textbf{Summary.} Surface the assumption a friendly read leaves buried by running an adversarial reviewer against every load-bearing claim before it is published.

\subsection*{Reasoning moves}
\begin{itemize}[leftmargin=1.5em]
  \item ask what would have to be true for the claim to fail
  \item run the adversarial reviewer before the friendly one
  \item treat an unaddressed objection as a publication blocker
\end{itemize}
\subsection*{Assumptions}
\begin{itemize}[leftmargin=1.5em]
  \item the adversarial reviewer is not subject to the same blind spot
  \item objections are cheaper to surface than to discover post-hoc
  \item a buried assumption is load-bearing until shown otherwise
\end{itemize}
\section{Results}
The cluster's conclusions were stress-tested against
1 resolved forecast(s) from the firm's
calibration ledger. Each row below resolves to a database record.

\begin{center}
\begin{tabular}{@{}p{0.30\textwidth} p{0.10\textwidth} p{0.12\textwidth} p{0.12\textwidth} p{0.20\textwidth}@{}}
\toprule
\textbf{Headline} & \textbf{Outcome} & \textbf{Stated} & \textbf{Brier} & \textbf{Row} \\
\midrule
Adversarial probing of hidden assumptions: claim 1 holds & YES & 0.70 & 0.090 & \rowref{forecast\_resolution:prediction-adversarial-audit-1} \\
\bottomrule
\end{tabular}
\end{center}

Across the 1 fully-scored forecast(s) above, the firm achieved a mean Brier score of 0.090. Each contributing row is auditable via its \textbackslash{}rowref\{...\} marker.
\section{Discussion}
Reviewer firm-peer-review on conclusion-adversarial-audit-1: verdict revise (confidence 0.78).
\subsection*{Peer-review findings}
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{minor} --- One supporting signal rests on a single source; triangulate before widening the claim.
        \rowref{review\_report:review-adversarial-audit-seed}
\end{itemize}
\section{Limits and failure modes}
The methodology root carries a registered failure-mode catalog. Every
limit below is an entry from the firm's catalog, not a synthesized
observation.

\begin{itemize}[leftmargin=1.5em]
  \item adversarial monoculture shares the friendly reviewer's blind spot
  \item objection fatigue downgrades a structural objection to a nitpick
\end{itemize}
\section{References}
\begin{enumerate}[leftmargin=2em]
  \item Firm source memo: Adversarial probing of hidden assumptions, May 2026 \rowref{artifact:artifact-adversarial-audit-source-memo}
\end{enumerate}

\vfill
\begin{center}\small
\textit{This paper was machine-drafted by Theseus from the firm's existing
artifacts. Every numerical claim resolves to a row in the firm's
database (see \rowref{...} markers). The .tex source is the authoritative
artifact; the PDF is a build product.}
\end{center}

\end{document}
