\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{titlesec}
\usepackage[hidelinks,colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black]{hyperref}
\usepackage{xcolor}
\usepackage{fancyhdr}

\titleformat{\section}{\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries}{\thesubsection}{1em}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Theseus Auto-Paper --- representational-geometry-109123c7}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\setlength{\parindent}{0pt}

\newcommand{\todomark}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\rowref}[1]{\texttt{\small[row:#1]}}

\title{\textbf{Geometric contradiction detection: a firm cluster}\\[6pt]\large Cluster representational-geometry-109123c7}
\date{May 2026}
\author{Theseus Auto-Paper Generator}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{center}
\fbox{%
  \begin{minipage}{0.92\textwidth}
    \centering\small
    \textbf{Disclosure:} machine-drafted, founder-reviewed.\\
    Lead conclusion: \texttt{conclusion-representational-geometry-1}.\\
    Methodology root: \texttt{profile-representational-geometry-1}
    (representational-geometry).
  \end{minipage}%
}
\end{center}

\vspace{1.5em}

\section*{Abstract}
This paper distills 2 firm conclusion(s) anchored on conclusion-representational-geometry-1, all sharing the methodology root "Geometric contradiction detection". The cluster's claims are stress-tested against 1 resolved forecast(s) drawn from the firm's calibration ledger. Lead claim: The geometry of a claim reveals a contradiction before the semantics of the claim does.

\section{Introduction}
The Theseus engine routinely produces firm conclusions, attaches them to a methodology profile, and tracks how those conclusions perform when their predictive consequences are settled in public forecast markets. This paper presents one such cluster of 2 conclusion(s), drawn together by their shared methodology root and by at least one resolved forecast that touches the cluster. The cluster's lead claim states: "The geometry of a claim reveals a contradiction before the semantics of the claim does.".

\subsection*{Conclusions in this cluster}
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{conclusion-representational-geometry-1} \rowref{conclusion:conclusion-representational-geometry-1} --- The geometry of a claim reveals a contradiction before the semantics of the claim does.
  \item \textbf{conclusion-representational-geometry-2} \rowref{conclusion:conclusion-representational-geometry-2} --- A contradiction shows up in the angle between a claim and its negation before it shows up in a close read.
\end{itemize}

\section{Methods}
The cluster shares a single methodology root,
\textit{Geometric contradiction detection} (\texttt{profile-representational-geometry-1}, pattern \emph{representational-geometry}).

\textbf{Summary.} Read the geometry of a claim's embedding to catch a contradiction before a semantic reading would.

\subsection*{Reasoning moves}
\begin{itemize}[leftmargin=1.5em]
  \item embed the claim and its negation
  \item measure the angle before reading the sentences
\end{itemize}
\subsection*{Assumptions}
\begin{itemize}[leftmargin=1.5em]
  \item the embedding geometry tracks semantic contradiction
\end{itemize}
\section{Results}
The cluster's conclusions were stress-tested against
1 resolved forecast(s) from the firm's
calibration ledger. Each row below resolves to a database record.

\begin{center}
\begin{tabular}{@{}p{0.30\textwidth} p{0.10\textwidth} p{0.12\textwidth} p{0.12\textwidth} p{0.20\textwidth}@{}}
\toprule
\textbf{Headline} & \textbf{Outcome} & \textbf{Stated} & \textbf{Brier} & \textbf{Row} \\
\midrule
Geometric contradiction detection: claim 1 holds & YES & 0.70 & 0.090 & \rowref{forecast\_resolution:prediction-representational-geometry-1} \\
\bottomrule
\end{tabular}
\end{center}

Across the 1 fully-scored forecast(s) above, the firm achieved a mean Brier score of 0.090. Each contributing row is auditable via its \textbackslash{}rowref\{...\} marker.
\section{Discussion}
Reviewer firm-peer-review on conclusion-representational-geometry-1: verdict revise (confidence 0.78).
\subsection*{Peer-review findings}
\begin{itemize}[leftmargin=1.5em]
  \item \textbf{minor} --- One supporting signal rests on a single source; triangulate before widening the claim.
        \rowref{review\_report:review-representational-geometry-seed}
\end{itemize}
\section{Limits and failure modes}
\todomark{methodology root has no registered failure modes; manual review required before publication}
\section{References}
\begin{enumerate}[leftmargin=2em]
  \item Firm source memo: Geometric contradiction detection, May 2026 \rowref{artifact:artifact-representational-geometry-source-memo}
\end{enumerate}

\vfill
\begin{center}\small
\textit{This paper was machine-drafted by Theseus from the firm's existing
artifacts. Every numerical claim resolves to a row in the firm's
database (see \rowref{...} markers). The .tex source is the authoritative
artifact; the PDF is a build product.}
\end{center}

\end{document}
