\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{xcolor}

\newcommand{\sP}{s_{\mathrm{P}}}
\newcommand{\sS}{s_{\mathrm{S}}}
\newcommand{\sAMF}{s_{\mathrm{AMF}}}
\newcommand{\sC}{s_{\mathrm{C}}}
\newcommand{\sDS}{s_{\mathrm{DS}}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\clip}{\operatorname{clamp}}

\title{Methodology Quality Score (MQS)\\\large Formal Specification, \texttt{SPEC\_VERSION} = 1.0.0}
\author{Theseus / Noosphere}
\date{2026-05-14}

\begin{document}
\maketitle

\begin{abstract}
The Methodology Quality Score is the operational form of the five working
criteria from \texttt{THE\_META\_METHOD.md}, attached to every Conclusion that
carries at least one MethodologyProfile. This is the first \emph{formal}
version of the specification: it states the closed form of each of the five
sub-scores, the exact gating function, the exact composite operator, and the
boundary cases. Equations are numbered and the body refers to them. The
specification is the source of truth for \code{noosphere/noosphere/evaluation/mqs.py};
\code{scripts/check\_mqs\_doc\_consistency.py} fails CI if the two drift. The
Markdown rendering at \code{docs/methods/MQS\_Specification.md} is the same
specification in prose; this PDF restates it in standard mathematical
notation. No equation in this specification calls an LLM judge --- the judge
produces sub-score \emph{evidence} and two numeric inputs, and the formulas
are arithmetic on numbers.
\end{abstract}

\section{Notation}
\label{sec:notation}

A conclusion $c$ carries text, a rationale, a topic hint, and links to
forecasts, dissent claims, and a domain bound. $M(c)$ is the set of
MethodologyProfiles attached to $c$ (possibly empty); $m \in M(c)$ is one
method. The five sub-scores and the composite are all in $[0,1]$:
\[
\sP,\ \sS,\ \sAMF,\ \sC,\ \sDS \in [0,1], \qquad \mathrm{MQS}(c) \in [0,1].
\]
$\sP$ is progressivity, $\sS$ severity, $\sAMF$ aim--method fit, $\sC$
compressibility, $\sDS$ domain sensitivity.

The typed inputs read from $c$ and $M(c)$:
\begin{itemize}
  \item $F(c) \in \mathbb{N}$ --- forecast count linked to $c$.
  \item $B(c) \in \{0,1\}$ --- $1$ iff $c$ carries a future check-back date.
  \item $R(c) \in \mathbb{N}$ --- decision-rule phrase count: regex matches of
        the decision-rule pattern in the conclusion text plus rationale.
  \item $\mathrm{FM}(c) = \bigcup_{m \in M(c)} \mathrm{failure\_modes}(m)$ ---
        concatenated failure-mode list; $A(c) = \bigcup_{m \in M(c)}
        \mathrm{assumptions}(m)$, with $n = |A(c)|$.
  \item $D(c) \in \mathbb{N}$ --- dissent-claim count.
  \item $j_{\mathrm{S}}, j_{\mathrm{DS}} \in [0,1] \cup \{\bot\}$ --- the LLM
        judge's numeric scores for severity and domain sensitivity; $\bot$
        means the judge produced no number.
  \item $\kappa \in [0,1] \cup \{\bot\}$ --- severity track-record ceiling.
  \item $\delta, \omega \in (0,1]$ --- severity drift penalty and objection
        penalty multipliers.
  \item $v \in \{\textsf{in\_bounds}, \textsf{edge\_case},
        \textsf{out\_of\_bounds}\} \cup \{\bot\}$ --- the domain-bound verdict.
\end{itemize}

$\clip(x)$ maps $x$ into $[0,1]$ and maps $\mathrm{NaN}$ to $0$. Every formula
in Section~\ref{sec:subscores} is \emph{total} --- defined for every input,
including $M(c) = \emptyset$ (Section~\ref{sec:boundary}).

\section{Sub-scores}
\label{sec:subscores}

\subsection{Progressivity $\sP$ --- deterministic}

Did the analysis produce a prediction, implication, or decision rule that can
later be checked? $\sP$ is a piecewise constant of $F(c), B(c), R(c)$; the
first matching clause wins:
\begin{equation}
\label{eq:progressivity}
\sP =
\begin{cases}
1.00 & B(c)=1 \ \wedge\ F(c)\ge 2 \ \wedge\ R(c)\ge 1,\\[2pt]
0.85 & \bigl(B(c)=1 \wedge F(c)\ge 1\bigr) \ \vee\ F(c)\ge 2,\\[2pt]
0.65 & B(c)=1 \ \vee\ F(c)\ge 1,\\[2pt]
0.40 & R(c)\ge 1,\\[2pt]
0.00 & \text{otherwise}.
\end{cases}
\end{equation}

\subsection{Severity $\sS$ --- deterministic floor, judged, then deterministic caps}

Would the procedure that produced $c$ have caught the claim if it were false?
$\sS$ is computed in five stages, each of which can only move the score toward
caution. The deterministic floor:
\begin{equation}
\label{eq:sev-floor}
\mathrm{floor} = \min\!\bigl(1,\ \texttt{SEVERITY\_FAILURE\_MODE\_FLOOR\_STEP}\cdot|\mathrm{FM}(c)|
 + \texttt{SEVERITY\_DISSENT\_FLOOR\_STEP}\cdot D(c)\bigr).
\end{equation}
The judge, floored (with $j_{\mathrm{S}} = \bot$ treated as $0$):
\begin{equation}
\label{eq:sev-judge}
\sS^{(0)} = \max\!\bigl(\mathrm{floor},\ \clip(j_{\mathrm{S}})\bigr).
\end{equation}
The no-failure-modes cap --- a method that lists no way it could fail is by
construction not severe:
\begin{equation}
\label{eq:sev-cap}
\sS^{(1)} =
\begin{cases}
\min\!\bigl(\sS^{(0)},\ \texttt{SEVERITY\_NO\_FAILURE\_MODES\_CAP}\bigr)
  & \mathrm{FM}(c)=\emptyset \ \wedge\ D(c)=0,\\[2pt]
\sS^{(0)} & \text{otherwise}.
\end{cases}
\end{equation}
The track-record ceiling --- a thin or poorly calibrated
\code{MethodTrackRecord} cannot project high severity onto new conclusions:
\begin{equation}
\label{eq:sev-ceiling}
\sS^{(2)} =
\begin{cases}
\min\!\bigl(\sS^{(1)},\ \clip(\kappa)\bigr) & \kappa \ne \bot,\\[2pt]
\sS^{(1)} & \kappa = \bot.
\end{cases}
\end{equation}
Finally the multiplicative penalties --- drift $\delta$ before objection
$\omega$, both in $(0,1]$:
\begin{equation}
\label{eq:sev-final}
\sS = \sS^{(2)} \cdot \delta \cdot \omega.
\end{equation}
Drift and objection penalties touch only severity: a method's recent
calibration and its open objections bound how far to trust \emph{this} claim
of having weighed counter-evidence; they do not touch the other criteria.

\subsection{Aim--method fit $\sAMF$ --- deterministic}

Is the method actually capable of answering the question being asked? Since
prompt~31 this sub-score is deterministic: it is the five-level rubric in
\code{noosphere/noosphere/inquiry/aim\_method\_fit.py}, driven by the question
typology. Let $\ell \in \{0,1,2,3,4\}$ be the rubric level assigned to $c$.
Then
\begin{equation}
\label{eq:amf}
\sAMF = \ell / 4 \in \{0.00, 0.25, 0.50, 0.75, 1.00\}.
\end{equation}
The rubric and its worked examples are specified in
\code{docs/methods/Aim\_Method\_Fit\_Rubric.md}. Aim--method fit asks whether
the method's \emph{output shape} matches the question; domain sensitivity
(Section~\ref{sec:domain}) asks whether the conclusion is inside the method's
\emph{domain}. The two are distinct: a method can be in-domain and still
misfit the question.

\subsection{Compressibility $\sC$ --- deterministic, LLM-pruned}

How many independent assumptions must hold for $c$ to survive? Let $n = |A(c)|$
and let $d \in [0,n]$ be the count of assumptions the LLM judge classifies as
decorative. The effective assumption count cannot fall below $1$:
\begin{equation}
\label{eq:comp-eff}
n' = \max(1,\ n - d).
\end{equation}
The score decays with each load-bearing assumption beyond the first:
\begin{equation}
\label{eq:comp}
\sC = \frac{1}{1 + (n' - 1)\cdot \texttt{COMPRESSIBILITY\_PENALTY\_STEP}}.
\end{equation}
With \texttt{COMPRESSIBILITY\_PENALTY\_STEP} $= 0.25$: $n' \in \{0,1\}
\mapsto 1.00$, $2 \mapsto 0.80$, $3 \mapsto 0.67$, $4 \mapsto 0.57$,
$5 \mapsto 0.50$. The LLM only ever reduces $n$; it cannot lift $\sC$ above
what the raw assumption count allows.

\subsection{Domain sensitivity $\sDS$ --- judged, with deterministic verdict override; the gate}
\label{sec:domain}

Where should this method stop being trusted, and is $c$ inside or outside that
domain? Define the pre-verdict score from the judge and a deterministic floor:
\begin{equation}
\label{eq:ds-base}
\sDS^{0} = \max(\mathrm{base},\ \mathrm{floor}),
\quad
\mathrm{base} =
\begin{cases}
\clip(j_{\mathrm{DS}}) & j_{\mathrm{DS}} \ne \bot,\\
\texttt{DOMAIN\_BACKFILL\_DEFAULT} & j_{\mathrm{DS}} = \bot,
\end{cases}
\quad
\mathrm{floor} =
\begin{cases}
\texttt{DOMAIN\_NO\_FAILURE\_MODES\_FLOOR} & \mathrm{FM}(c)=\emptyset,\\
0 & \text{otherwise}.
\end{cases}
\end{equation}
The domain-bound verdict $v$ then overrides:
\begin{equation}
\label{eq:ds}
\sDS =
\begin{cases}
0 & v = \textsf{out\_of\_bounds},\\[2pt]
\min\!\bigl(\sDS^{0},\ \texttt{EDGE\_CASE\_DOMAIN\_CEILING}\bigr) & v = \textsf{edge\_case},\\[2pt]
\sDS^{0} & \text{otherwise}.
\end{cases}
\end{equation}
An \textsf{out\_of\_bounds} verdict is a hard zero --- the orchestrator has
established $c$ is outside the method's declared \code{DomainBound}. The
deterministic floor \texttt{DOMAIN\_NO\_FAILURE\_MODES\_FLOOR} $= 0.10$ keeps a
method with no declared failure modes off zero --- its domain claim is
\emph{unverifiable}, not \emph{failed} --- but, crucially, that floor is below
the gate threshold $\theta$ (Section~\ref{sec:composite}), so the floor alone
never opens the gate.

\section{Composite}
\label{sec:composite}

Domain sensitivity is a \emph{gate}, not a weighted addend and --- since this
formal version --- not a soft multiplicative penalty either. With gate
threshold $\theta = \texttt{DS\_GATE\_THRESHOLD} = 0.15$:
\begin{equation}
\label{eq:composite}
\mathrm{MQS}(c) =
\begin{cases}
0 & \sDS < \theta,\\[3pt]
\operatorname{wgm}(\sP, \sS, \sAMF, \sC) & \sDS \ge \theta.
\end{cases}
\end{equation}
The gate is closed on the passing side: $\sDS = \theta$ exactly opens it. The
combining operator is the weighted geometric mean, with the
\code{SUBSCORE\_WEIGHTS} as exponents:
\begin{equation}
\label{eq:wgm}
\operatorname{wgm}(\sP, \sS, \sAMF, \sC)
 = \sP^{\,w_{\mathrm{P}}}\cdot \sS^{\,w_{\mathrm{S}}}\cdot \sAMF^{\,w_{\mathrm{AMF}}}\cdot \sC^{\,w_{\mathrm{C}}},
\qquad \sum w_i = 1.
\end{equation}
The four weights are equal at $0.25$, so $\operatorname{wgm} = (\sP\, \sS\,
\sAMF\, \sC)^{1/4}$. The operator has the weakest-link property:
\begin{equation}
\label{eq:weakest-link}
\bigl(\exists\, i:\ x_i = 0\bigr) \ \Longrightarrow\ \operatorname{wgm}(x) = 0.
\end{equation}
It is monotone non-decreasing in each argument, equals $1$ if and only if every
argument equals $1$, and stays in $[0,1]$.

\subsection{Why this composite}

The Round~17 prose draft used $\mathrm{MQS}(c) = \sDS \cdot
\operatorname{mean}(\sP, \sS, \sAMF, \sC)$ --- a soft penalty in which
$\sDS = 0.5$ merely capped the composite at $0.5$. The firm's position on a
method pointed outside its domain is stronger than a discount: such a method
is not producing a worse answer, it is producing one the firm has no basis to
trust. Hence the hard gate of Eq.~\eqref{eq:composite}. Among the three
operators that share the weakest-link property of
Eq.~\eqref{eq:weakest-link} --- the minimum, the harmonic mean, and the
weighted geometric mean --- the minimum discards three of four signals and the
harmonic mean is so dominated by the smallest input that the others barely
register. The weighted geometric mean keeps all four with proportional weight,
still refuses redemption, and is the exact multiplicative analogue of the
prose draft's arithmetic mean (the same weight map, used as exponents). It is
the choice.

\section{Boundary cases}
\label{sec:boundary}

Each rule below is pinned by a named test in
\code{noosphere/tests/test\_mqs\_spec.py}.

\paragraph{Insufficient data.} No sub-score is ever undefined. With
$M(c) = \emptyset$: $\sP$ is still computed from $c$;
Eq.~\eqref{eq:sev-floor} gives $\mathrm{floor}=0$ and Eq.~\eqref{eq:sev-cap}
caps $\sS$ at $0.35$; $\sAMF = 0.5$ (rubric level $2$, ``cannot verify fit'');
Eq.~\eqref{eq:comp-eff}--\eqref{eq:comp} give $n'=1$ and $\sC = 1.0$;
Eq.~\eqref{eq:ds-base} floors $\sDS$ at \texttt{DOMAIN\_NO\_FAILURE\_MODES\_FLOOR}
and, with no judge, uses $\mathrm{base} = \texttt{DOMAIN\_BACKFILL\_DEFAULT}$.
Insufficient data resolves to a defined, conservative number.

\paragraph{At the gate threshold.} The gate condition $\sDS < \theta$ is
strict, so $\sDS = \theta$ opens the gate and $\sDS = \theta - \varepsilon$
closes it. The composite is discontinuous at $\theta$: this is intentional ---
it makes the gate a verdict, not a dial. Note
$\texttt{DOMAIN\_NO\_FAILURE\_MODES\_FLOOR} = 0.10 < \theta = 0.15$.

\paragraph{Inputs disagree.} When a deterministic signal and the LLM judge
disagree, the deterministic signal wins, toward caution: an
\textsf{out\_of\_bounds} verdict forces $\sDS = 0$ over any $j_{\mathrm{DS}}$
(Eq.~\eqref{eq:ds}); the no-failure-modes cap overrides a high $j_{\mathrm{S}}$
(Eq.~\eqref{eq:sev-cap}); the deterministic floor overrides a low
$j_{\mathrm{S}}$ via the $\max$ of Eq.~\eqref{eq:sev-judge}; the track-record
and edge-case ceilings override high judge scores via $\min$. The judge moves
a sub-score only within the window the deterministic rules leave open.

\paragraph{A non-gate sub-score is exactly $0$.} If any of $\sP, \sS, \sAMF,
\sC$ is $0$ and the gate is open, Eq.~\eqref{eq:weakest-link} gives
$\mathrm{MQS}(c) = 0$. This is intentional: a conclusion that produces nothing
checkable, or whose method cannot answer the question at all, has a failing
methodology quality regardless of the other axes.

\paragraph{Out-of-range or NaN inputs.} Every input passes through $\clip$
before any arithmetic; $\clip$ confines values to $[0,1]$ and maps
$\mathrm{NaN}$ to $0$. The composite is always a real number in $[0,1]$.

\section{Composite tiers}

The composite is continuous in $[0,1]$, but the firm reasons about it in four
tiers, each an inclusive lower bound:
\begin{center}
\begin{tabular}{lc}
\toprule
Tier & Composite $\ge$\\
\midrule
\texttt{strong}      & 0.66\\
\texttt{adequate}    & 0.40\\
\texttt{provisional} & 0.15\\
\texttt{failing}     & 0.00\\
\bottomrule
\end{tabular}
\end{center}
\code{tier\_rank} gives each tier an ordinal ($\texttt{failing}=0,\dots,
\texttt{strong}=3$). A re-score that drops a conclusion a tier is routed to
the founder's queue rather than silently downgraded.

\section{Constants registry}

Every constant the specification depends on, mirrored exactly by
\code{MQS\_CONSTANTS} in the code and checked bidirectionally by
\code{scripts/check\_mqs\_doc\_consistency.py}.
\begin{center}
\begin{longtable}{lll}
\toprule
Constant & Value & Role\\
\midrule
\endhead
\texttt{SPEC\_VERSION} & 1.0.0 & version of this specification.\\
\texttt{MQS\_SCHEMA} & theseus.mqs.v1 & schema string on every MQS row.\\
\texttt{PROMPT\_VERSION} & mqs-prompt-v2.0 & LLM-judge prompt version.\\
\texttt{DS\_GATE\_THRESHOLD} & 0.15 & gate threshold $\theta$ (Eq.~\eqref{eq:composite}).\\
\texttt{COMPOSITE\_OPERATOR} & weighted\_geometric\_mean & combining operator (Eq.~\eqref{eq:wgm}).\\
\texttt{SEVERITY\_FAILURE\_MODE\_FLOOR\_STEP} & 0.15 & per-failure-mode floor step (Eq.~\eqref{eq:sev-floor}).\\
\texttt{SEVERITY\_DISSENT\_FLOOR\_STEP} & 0.10 & per-dissent floor step (Eq.~\eqref{eq:sev-floor}).\\
\texttt{SEVERITY\_NO\_FAILURE\_MODES\_CAP} & 0.35 & cap when no failure modes (Eq.~\eqref{eq:sev-cap}).\\
\texttt{COMPRESSIBILITY\_PENALTY\_STEP} & 0.25 & per-assumption penalty step (Eq.~\eqref{eq:comp}).\\
\texttt{DOMAIN\_NO\_FAILURE\_MODES\_FLOOR} & 0.10 & $\sDS$ floor when no failure modes (Eq.~\eqref{eq:ds-base}).\\
\texttt{DOMAIN\_BACKFILL\_DEFAULT} & 0.5 & $\sDS$ base when no judge (Eq.~\eqref{eq:ds-base}).\\
\texttt{EDGE\_CASE\_DOMAIN\_CEILING} & 0.4 & $\sDS$ cap under \textsf{edge\_case} (Eq.~\eqref{eq:ds}).\\
\texttt{EVIDENCE\_STR\_CAP} & 600 & per-string evidence-blob character cap.\\
\bottomrule
\end{longtable}
\end{center}
The four \code{SUBSCORE\_WEIGHTS} (all $0.25$) and the \code{COMPOSITE\_TIERS}
lower bounds are checked separately by the same script.

\section{Versioning and changelog}

\texttt{SPEC\_VERSION} versions this specification; the first formal version
is \texttt{1.0.0}. Any material change to a formula, a constant, a boundary
rule, or the composite operator bumps \texttt{SPEC\_VERSION} and adds a
changelog row. \texttt{PROMPT\_VERSION} versions the LLM-judge prompt and is
independent.

\begin{center}
\begin{tabular}{p{0.10\textwidth}p{0.13\textwidth}p{0.67\textwidth}}
\toprule
\texttt{SPEC\_VERSION} & Date & Change\\
\midrule
1.0.0 & 2026-05-14 & First formal specification. The composite changed from
the Round~17 prose draft's soft multiplicative gate, $\sDS \cdot
\operatorname{mean}(\sP,\sS,\sAMF,\sC)$, to the hard piecewise gate of
Eq.~\eqref{eq:composite} at $\theta = 0.15$ followed by the weighted geometric
mean of Eq.~\eqref{eq:wgm}. Both the code (\code{composite\_score}) and the
spec were changed together --- they had previously agreed on the soft formula.
Rationale in Section~\ref{sec:composite}. Every previously-magic constant was
named and the constants registry added.\\
\bottomrule
\end{tabular}
\end{center}

\end{document}
